---
title: "Create fishery diversification indices for CEFI analysis"
author: Abigail Golden
output: html_document
date: "2025-01-03"

  html_document:
    df_print: paged
    toc: yes
    toc_depth: '5'
  pdf_document:
    highlight: haddock
    number_sections: yes
    toc: yes
    toc_depth: '3'
geometry: margin=1in
fontsize: 11pt
---

# Description

This document generates fishery diversification indices for US West Coast port groups based on vessel-level landings and revenue from fish ticket data.

# Setup

<br>
```{r "setup", include=FALSE}
if(!require("here")) {install.packages("here")}
library(here)
knitr::opts_chunk$set(echo = TRUE)
knitr::opts_knit$set(root.dir = here::here())

## start time for full script
script_start_time <- Sys.time()
```

<br>

This script requires the following packages. 
```{r packages, message=FALSE, warning=FALSE}
if(!require("tidyverse")) {install.packages("tidyverse")}
if(!require("lubridate")) {install.packages("lubridate")}
if(!require("ggplot2")) {install.packages("ggplot2")}
```

And calls the following functions:
```{r functions}
source(here::here("R/generate_fishery_matrix.R")) # creates matrix of revenue by species group x vessel 
source(here::here("R/calculate_esi.R")) # creates fishery diversification indices

```

# User Inputs 

Select your directories.
```{r directories}
## location of fish ticket with metiers assigned, metier key for each port group
indir = "/Users/abigail.golden/Desktop/Research/24 - CEFI/data/fish_tickets/processed"

## output directory for igraph objects
outdir_ports = 'data/networks/participation_vessel_ports'

# outdir_states = 'data/networks/participation_vessel_states'

## output directory for network viz
pngdir_ports = 'data/networks/participation_vessel_ports/plots'
# pngdir_states = 'data/networks/participation_vessel_states/plots'

## output file (including directory) for stat summaries
statdir = 'results/statistics'

## directory for species group key. not confidential
processdir <- "data/input"
```

<br>

Specify file name of the species groupings lookup key that contains labels for the network graphs.
```{r get_filenames}

myfile5 <- "spgrpn2_iea_key.csv"

```
<br>

Identify the years and port groups for which you would like to calculate diversification indices.

```{r}
ref_years <- c(2004,2005)

## crab years
years <- seq(2005, 2024)

## IOPAC port groups
myports <- c("Puget Sound","North WA Coast","WA Coast","Other Coastal Washington","Astoria","Tillamook","Columbia River","Newport","Coos Bay","Brookings","Crescent City","Eureka","Morro Bay","Fort Bragg","Bodega Bay","San Francisco","San Diego","Monterey","Santa Barbara","Los Angeles","Unknown Ports" )

wa_ports <- myports[1:4]
or_ports <- myports[5:10]
ca_ports <- myports[11:20]

## west coast states
mystates <- c("C", "O","W")

```

For confidentiality, three or more vessels must be participating in a given fishery for that fishery to be included in the networks. To avoid inclusion of rare or minimal fishing activity, a given fishery must contribute to at least 10% of a vessel's seasonal revenue to be included in the network data. 

The cutoff values can be increased if more stringent fishery inclusion criteria are preferred, and the contribution cutoff may be decreased (i.e. for summaries over shorter temporal / smaller spatial scales).
```{r}

vessel_cutoff <- 3
contr_cutoff <- 0.10

```

# Read in the data

Read in the landings data, from the file containing all fish tickets across port groups, within a single year.

NOTE: this step takes a while.

NOTE: read in fish ticket files with this naming convention: fish_tickets_crab_2004_processed_for_networks.csv

```{r}

# date of last fish ticket processing
rundate <- "2025-01-08"

for(y in years){
  if(y %in% ref_years){
    tmptix <- read.csv(paste0(indir,"/fish_tickets_crab_", y, "_processed_for_networks_",rundate, ".csv"), stringsAsFactors = FALSE) %>%
    filter(IOPAC %in% myports)
  } else{
    tmptix <- read.csv(paste0(indir,"/fish_tickets_crab_", y, "_processed_for_networks_",rundate, ".csv"), stringsAsFactors = FALSE) %>%
    filter(IOPAC %in% myports)
  }
  if(exists('fishtix')){
    fishtix <- rbind(fishtix, tmptix)
  } else{ fishtix <- tmptix }
}
rm(tmptix)

# some checks on species groups
sort(unique(fishtix$SPGRPN2))
length(which(is.na(fishtix$SPGRPN2)))
unique(fishtix[which(is.na(fishtix$SPGRPN2)),]$spid)
unique(fishtix[which(is.na(fishtix$SPGRPN2)),]$year)
unique(fishtix[which(is.na(fishtix$SPGRPN2)),]$agid)
unique(fishtix[which(is.na(fishtix$SPGRPN2)),]$pcid)

```

Make sure that all dates are `Posixt` objects and species group labels are characters.
```{r}
#dates_df$odate <- mdy(dates_df$odate)
fishtix$tdate <- date(parse_date_time(fishtix$tdate, orders=c("ymd", "mdy")))
fishtix$SPGRPN2 = as.character(fishtix$SPGRPN2)
```
<br>

Species group names
```{r}

#### Read in species group labels key ####
# formerly this object was named `node_names`
species_grp_key <- read_csv(here::here(processdir,myfile5),
                       col_types = list(col_character(), col_character(), col_character()))

```
<br>

# Calculate fishery diversification indices

We only include vessels that generate >=`$5,000` in total fisheries revenue each year and >=`$500` in revenue from individual fisheries.

```{r cutoffs}

total_rev_cutoff <- 5000
indiv_rev_cutoff <- 500

```
<br>

Generate matrix of revenue by species group
```{r}

mat <- revenue_matrix(tickets = fishtix, matrix_scale = "coastwide", matrix_type = "week", pcid_choose = NA, year_choose = 2019,
                      min_rev = total_rev_cutoff, min_rev_indiv = indiv_rev_cutoff)

```


Generate fishery diversification index for all individual vessels across all ports across all years


```{r}

for (y in years){
  for (p in myports){
    cat(p, "\n")
    
    # generate revenue matrix
    
    mat <- revenue_matrix(tickets = fishtix, matrix_scale = "port", matrix_type = "species", 
                          pcid_choose = p, year_choose = y,
                          min_rev = total_rev_cutoff, min_rev_indiv = indiv_rev_cutoff)
    
    # calculate esi
    
    tmp_esi <- calculate_esi(mat = mat)
    
    tmp_esi$port <- p
    tmp_esi$year <- y
    
    if(exists("spp_divers_index")){
      spp_divers_index <- rbind(spp_divers_index, tmp_esi)
    } else {
      spp_divers_index <- tmp_esi
    }
    
  }
}

```

