---
title: "Sensitivity Index code"
output: html_document
date: "2024-12-31"
author: "Amanda Phillips"
---



```{r include=FALSE}
library(tidyverse)
library(ggthemes)
library(cowplot)
library(viridis)
library(kableExtra) # create data tables
library(scales) #commas for axes in ggplot 
library(ggpubr) #ggarrange()
library(ggrepel)
library(vegan)
library(zoo)
library(fmsb)
library(paletteer)
library(GGally)
library(nord)

```

```{r include = FALSE}
knitr::opts_chunk$set(fig.path='./Figures/', echo= F, warning= F, message= F, dev = 'png')
options(ggrepel.max.overlaps = Inf)
extrafont::loadfonts(device="win")

theme_set(theme_tufte())
```

1. Read in landings data and iopac data
- Data manipulation:
 -- Set revenue to millions
 -- Restrict included years
 -- convert 0's to 0.000000000001 for division purposes. 

```{r include=FALSE}

# directories
out_conf_dir <- "/Users/abigail.golden/Desktop/Research/24 - CEFI/data/fish_tickets/processed/"
processdir <- "data/input"

# file names
myfile2 <- 'fish tickets 2004-2024 121724.rds' # latest pull from PacFIN 12-09-2024
myfile3 <- "iopac_conversion_table.csv"

#read in landings data
rawdat <- read_rds(paste0(out_conf_dir,myfile2))

# read in iopac data
port_grps <- read_csv(here::here(processdir,myfile3))


# landings_iopac_mgrp <- readRDS(file = "./Data/landings_iopac_mgrp_crabyr.RDS")
db <- rawdat %>%
  mutate(AFI_REV_m = AFI_EXVESSEL_REVENUE/1000000)  # convert landings to millions
 # mutate(IOPAC_PORT_GROUP = case_when(IOPAC_PORT_GROUP == 'SOUTH AND CENTRAL WA COAST' ~ 'WA COAST',
                                                          #      T ~ IOPAC_PORT_GROUP))

# db[db==0] <- .000000000001 # convert 0's to near 0's for index calculation

```

2. Join geographic data to landings data (IOPAC + landings)


```{r include=FALSE}


# read in iopac data
#iopac <- read_csv('./Data/iopac_lat-long-r.csv')
#iopac$IOPAC_PORT_GROUP <- toupper(iopac$IOPAC_PORT_GROUP)



db.1 <- db %>% 
  left_join(port_grps) %>% 
  #filter(IOPAC != "Unknown Ports") %>% #filter out tickets not associated with an IOPAC port
  arrange(desc(PACFIN_LAT)) %>% # arrange latitudinally
  mutate(IOPAC_PORT_GROUP = factor(IOPAC_PORT_GROUP, levels = unique(IOPAC_PORT_GROUP))) # make IOPAC a factor variable

# sum_revenue <- sum(db.1$AFI_REV_m, na.rm = TRUE)
# 
# db.1 %>% 
#   group_by(IOPAC) %>% 
#   summarize(pct_revenue = sum(AFI_REV_m, na.rm = TRUE)/sum_revenue*100)

# db <- db %>% left_join(port_grps) %>%
#   filter(IOPAC_PORT_GROUP != 'NOT_DESIGNATED') %>%
#   arrange(desc(PRIMARY_LATITUDE)) %>%
#   mutate(IOPAC_PORT_GROUP = factor(IOPAC_PORT_GROUP, levels = unique(IOPAC_PORT_GROUP)))

db_ma <- db.1 %>% 
  #filter(MANAGEMENT_GROUP_CODE == 'TOTAL') %>% # this is a species group code, I think one that AP and others created to summarize revenue across all PacFIN species
    group_by(IOPAC_PORT_GROUP) %>%
  mutate(rm_REV = rollmean(AFI_REV_m, 3, fill = NA, align = "right")) # calculate rolling mean for previous three years 

db_full <- db.1 %>%
  filter(!IOPAC_PORT_GROUP %in% c('NOT_DESIGNATED', 'COASTWIDE') & MANAGEMENT_GROUP_CODE != 'TOTAL') %>%
  group_by(IOPAC_PORT_GROUP, CRAB_YEAR, PRIMARY_LATITUDE) %>% 
  mutate(total_revenue = sum(AFI_REV_m)) %>% # calculate total revenue by port and year
  ungroup() %>% 
  group_by(IOPAC_PORT_GROUP, CRAB_YEAR, MANAGEMENT_GROUP_CODE, PRIMARY_LATITUDE) %>% 
  mutate(mngmt_revenue = sum(AFI_REV_m, na.rm = TRUE), # calculate total revenue by year, port and management group 
         percent_total = mngmt_revenue/total_revenue)  %>% 
  ungroup() %>% 
  group_by(IOPAC_PORT_GROUP, MANAGEMENT_GROUP_CODE)


lats <- iopac %>% select(IOPAC_PORT_GROUP, PRIMARY_LATITUDE) # create table with latitude information for future table

# region <- db_full %>% ungroup() %>% select(IOPAC_PORT_GROUP, REGION)

```

```{r}

fisheries_summary <- db %>% filter(CRAB_YEAR == 2019) %>% dplyr::group_by(MANAGEMENT_GROUP_CODE) %>% dplyr::summarize(ann_rev = sum(AFI_REV_m)) %>% 
  mutate(MANAGEMENT_GROUP_CODE = case_when(MANAGEMENT_GROUP_CODE == 'TOTAL' ~ 'AFSH', T ~ MANAGEMENT_GROUP_CODE))

fs <- fisheries_summary %>% left_join(mc) 

```

```{r, fig.height = 6, fig.width = 8}

fisheries_summary <- db %>% filter(CRAB_YEAR >= 2000 & CRAB_YEAR <= 2012) %>% dplyr::group_by(MANAGEMENT_GROUP_CODE, CRAB_YEAR) %>% dplyr::summarize(ann_rev = sum(AFI_REV_m)) %>% filter(MANAGEMENT_GROUP_CODE != 'TOTAL')

fs <- fisheries_summary %>% left_join(mc) 
k <- nord("algoma_forest", 7, reverse = T)

fs_plot <- ggplot(fs, aes(x=CRAB_YEAR,y=ann_rev,color=MANAGEMENT_NAME_COMBINED)) + 
  annotate("rect",xmin=2007,xmax=2008,ymin=0,ymax=Inf,alpha=0.2,fill="#4c556a") +
  annotate("rect",xmin=2005,xmax=2007,ymin=0,ymax=Inf,alpha=0.2,fill="#D8DEE9") + 
  geom_line(size = 1.25) + 
  geom_point(size = 2) +
  geom_text_repel(seed = 42, data = subset(fs, CRAB_YEAR == 2012), aes(x=CRAB_YEAR, y = ann_rev, label= MANAGEMENT_NAME_COMBINED), size = 2.75, nudge_x = 1, show.legend = FALSE, xlim  = c(2012, NA)) +
  scale_color_discrete(name="Management Group Code") + 
  theme_tufte(ticks = T) +
  theme(legend.position = "none") + 
  ylab("Real Revenue (millions of $s)") +
  xlab("") + 
  scale_color_manual(values = k) +
  scale_x_continuous(breaks=c(2000:2012), expand = expansion(mult = c(0, 0.3)))

fs_plot


```


```{r include = FALSE}
# create a variety of vectors for labeling tables and figures

variables <- c('AFI_REV_m', 'NUM_VESSELS', 'NUM_DEALERS')

iopac_ports <- unique(db$IOPAC_PORT_GROUP) 
region <- unique(db$REGION)
mgmt <- unique(db$MANAGEMENT_GROUP_CODE)

shock <-  c('E1', 'E2', 'E3', 'E4', 'E5', 'M1', 'M3', 'M2') # shorthand code for the different shocks identified in the paper
category <-  c('Ecological', 'Ecological', 'Ecological', 'Ecological', 'Ecological',  
             'Management', 'Management', 'Management') # Type of shock
type <-  c('El Nino', 'El Nino', 'El Nino', 'El Nino', 'Marine Heat Wave', 
         'Vessel Buyback', 'Catch Share Implementation', 'Southern California Current Salmon Closures') # specific shock long title
duration <-  c('1986 - 1988', '1991 - 1992', '1997 - 1998', '1998 - 1999', '2015 -2016', 
             '2003 - 2004', '2011 - 2012', '2008 - 2010') # time range for shocks
first_year <-  c(1986, 1991, 1997, 1998, 2015, 2003, 2011, 2008) # first year for each shock
end_year <-  c(1988, 1992, 1998, 1999, 2016, 2004, 2012, 2010) # last year for each shock 

shock_table <- data.frame(Shock_id = shock, 
                          Category = category,
                          Shock = type,
                          Duration = duration,
                          Start_year = first_year,
                          End_year = end_year) %>% 
   arrange(Shock_id) # create a table with the metadata for all shock scenarios, combining above vectors 

shock_table

```

```{r}
stk <- shock_table %>% select(Shock, Duration) %>% 
  kable() %>% 
  pack_rows(
  index = c("Ecological" = 5, "Management" = 3)) %>% 
  row_spec(0, font_size = 20) %>% 
  kable_classic(full_width = F, html_font = "Cambria") 

  # save_kable(file = "shock_table.png", self_contained = T)

stk
```
## Sensitivity Calculations: 

* Year prior to shock date uses average of 3 years prior to shock beginning to calculate sensitivity index
* value less than -1 is highly sensitive, value greater than -1 is low sensitivity. 


The year prior value for this sheet is the year prior to the first year listed in the shock scenario: 

* for the 1991-1992 El Nino shock:
year prior = average of 1990, 1989, 1988 revenue values
shock year = 1992 revenue values

```{r include=FALSE}
# create function to calculate the index for a given shock - takes the label "shock" as determined in the shock table listed above 

PCDiD_index <- function(shock) {
  st <- shock_table %>% filter(Shock_id == shock)
  year_prior <- st$Start_year - 1 # designate year prior to the shock 
  
  py_rev <- db_ma %>% filter(CRAB_YEAR == year_prior) %>% select(IOPAC_PORT_GROUP, rm_REV) # create prior year datatable with only data from the year prior to the shock, with port group and rolling mean for revenue 
  sy_rev <- db_ma %>% filter(CRAB_YEAR == st$Start_year) %>% select(IOPAC_PORT_GROUP, AFI_REV_m) # create datatable with only data from the start year of the shock, using revenue from that year (not rolling mean)
  
  rev_tbl <- merge(py_rev, sy_rev) # combine the prior year table and start year table 
  
  cw_yearprior <- rev_tbl[rev_tbl$IOPAC_PORT_GROUP == 'COASTWIDE', 'rm_REV'] # isolate the coastwide revenue for prior years
  cw_yearshock <- rev_tbl[rev_tbl$IOPAC_PORT_GROUP == 'COASTWIDE', 'AFI_REV_m'] # isolate the coastwide revenue for start year
  
  
  si_tbl <- rev_tbl %>% 
  mutate(index_type = 'PCDiD',
         port_diff = (AFI_REV_m - rm_REV), # calculate the difference between the prior rolling mean revenue and start year mean revenue for each port
         change_in_port = (AFI_REV_m - rm_REV) / rm_REV, # calculate the change in difference between start year mean revenue and rolling mean in port
         cw_diff = cw_yearshock - cw_yearprior, # calculate the coastwide difference in revenue for prior years and start year
         change_in_cw = (cw_yearshock - cw_yearprior) / cw_yearprior,  # calculate the change in coastwide difference for start year mean revenue and prior year revenue
         sensitivity_index = (AFI_REV_m/cw_yearshock) - (rm_REV/cw_yearprior), # calculate the sensitivity index using the port start revenue/coastwide start year revenue - port prior year revenue/coastwide prior year revenue
         Shock_id = shock) %>% 
    left_join(st) %>% 
    mutate(shock_description = paste(Duration, ": ", Shock))
  
  return(si_tbl)
} 


```

```{r include=FALSE}

index_table <- bind_rows(lapply(shock, PCDiD_index)) %>% # apply the sensitivity function for all shock scenarios
  mutate(port_change_description = case_when(port_diff > 1 ~ "revenue gain", 
                                             port_diff < 1 ~ "revenue loss", 
                                             port_diff == 1 ~ "no change")) %>% 
  left_join(lats)

# write_rds(index_table, './table_allindextypes.RDS')

si_ports <- index_table 
si_ports

```



```{r}


db_revenue <- db %>% filter(MANAGEMENT_GROUP_CODE != 'TOTAL') %>%  
  select(CRAB_YEAR, IOPAC_PORT_GROUP, MANAGEMENT_GROUP_CODE, AFI_REV_m) # create df with revenue for each management group by port

db_revenue_w <- db_revenue %>% # create wide version of above table
  pivot_wider(names_from = MANAGEMENT_GROUP_CODE, values_from = AFI_REV_m)


```



```{r}

dbr <- db_revenue %>% group_by(MANAGEMENT_GROUP_CODE, IOPAC_PORT_GROUP) %>% # prepare table to create shannon index
  mutate(rm_REV = rollmean(AFI_REV_m, 3, fill = NA, align = "right")) %>% 
  select(CRAB_YEAR, IOPAC_PORT_GROUP, MANAGEMENT_GROUP_CODE, rm_REV) %>% 
  pivot_wider(names_from = MANAGEMENT_GROUP_CODE, values_from = rm_REV) %>% 
  filter(!is.na(CPEL))

shannon_s <- ddply(dbr, CRAB_YEAR~IOPAC_PORT_GROUP,function(x) { # calculate shannon index
     data.frame(SHANNON_revenue=diversity(x[-2], index="shannon")) }) 

shannon_db <- shannon_s %>% # clean up shannon table
  left_join(lats) %>% 
  arrange(PRIMARY_LATITUDE) %>% 
  mutate(IOPAC_PORT_GROUP = factor(IOPAC_PORT_GROUP, levels = unique(IOPAC_PORT_GROUP)))


# saveRDS(shannon_db, './data/shannon_db.RDS')
```


## Shannon Index
 
 * x-axis: shannon diversity index for the year n-1 of the shock, where shannon indices are calculated using 3 year rolling averages
 * y-axis: sensitivity index 
 * each point represents a single port with matching shannon index and sensitivity index for a given shock

 

```{r}
region_db <- db %>% ungroup() %>% select(IOPAC_PORT_GROUP, REGION) %>% distinct()


yr <- si_ports %>% 
  select(Start_year, End_year, shock_description, Category) %>% 
  distinct() %>%
  mutate(Length = 1.5, year_prior = Start_year -1.5, year_post = End_year + .5, year_mid = (Start_year + End_year)/2) 

sh_sens <- si_ports %>% filter(IOPAC_PORT_GROUP != 'COASTWIDE') %>% 
    mutate(Year_prior = Start_year) %>% 
    left_join(shannon_db, by = c(c(IOPAC_PORT_GROUP = 'IOPAC_PORT_GROUP'), c(Start_year = 'CRAB_YEAR'))) %>% 
    left_join(region_db) 

sh_sens$REGION <- factor(sh_sens$REGION, level = c('Washington', 'Oregon', 'Northern California', 'Southern California'))

avg_sh_c <- mean(sh_sens$SHANNON_revenue, na.rm = TRUE) + sd(sh_sens$SHANNON_revenue, na.rm = TRUE)
# avg_sh <- 1.59


```

# Figure for shannon Index

```{r, shannonrev_ts, fig.height = 4, fig.width = 18}
scatter_plot_r2 <- sh_sens %>% ggplot() +
  # geom_rect(aes(xmin = 0, xmax = 0.4, ymin = -1, ymax = 10), fill = 'green', alpha = 0.05) +
  geom_point(aes(x = SHANNON_revenue, y = sensitivity_index)) +
  geom_hline(yintercept = 0 ) +
  geom_smooth(aes(x = SHANNON_revenue, y = sensitivity_index, color = Category, fill = Category)) +
  ggtitle('Shannon Index of Diversity and Sensitivy Index by Shock') +
  facet_grid(cols = vars(Shock)) +
    xlab("Shannon Index") +
  scale_color_manual(values=c('Ecological' = "#8D9E7C", 'Management' = '#9EAFBF'), name = "") +
  scale_fill_manual(values=c('Ecological' = "#8D9E7C", 'Management' = '#9EAFBF'), name = "") +
  ylab("Sensitivity Index")
#

scatter_plot_r2
```


## Visualizing sensitivity based on shocks
* These figures all attempt to gauge the range of sensitivity for a given shock



## Lollipop chart - by shock


**fig 3.2 sensitivity for port by shock -lollipop**

```{r sensitivity_comb, fig.height = 12, fig.width = 14}

vp_df <- si_ports %>% dplyr::select(-shock_description) 

si_ports <- si_ports %>% arrange(PRIMARY_LATITUDE) %>% 
  mutate(IOPAC_PORT_GROUP = factor(IOPAC_PORT_GROUP, levels = unique(IOPAC_PORT_GROUP)))

si_ports <- si_ports %>% mutate(IOPAC_PORT_GROUP_LC = str_to_title(IOPAC_PORT_GROUP),
                                IOPAC_PORT_GROUP_TT = case_when(IOPAC_PORT_GROUP_LC == "Wa Coast" ~ "WA Coast",
                                                                IOPAC_PORT_GROUP_LC == "North Wa Coast" ~ "North WA Coast",
                                                                T ~ IOPAC_PORT_GROUP_LC)) %>%  # rename port groups for figures 
  arrange(PRIMARY_LATITUDE) %>% 
  mutate(IOPAC_PORT_GROUP_TT = factor(IOPAC_PORT_GROUP_TT, levels = unique(IOPAC_PORT_GROUP_TT))) %>% filter(IOPAC_PORT_GROUP_TT != 'Coastwide')




SI_ll <- si_ports %>%  filter(Category == 'Ecological') %>% 
  filter(!IOPAC_PORT_GROUP %in% c('NOT_DESIGNATED')) %>% # filter out na's 
  ggplot(aes(x = sensitivity_index, y = IOPAC_PORT_GROUP_TT)) +
  geom_point(aes(color = port_change_description, fill = port_change_description), size=3, alpha=0.7, stroke = 2) +
  geom_segment(aes(y=IOPAC_PORT_GROUP_TT, yend=IOPAC_PORT_GROUP_TT, x=0, xend=sensitivity_index), color = "grey40", size = .75) +
  geom_vline(xintercept = 0) +
  facet_wrap(~shock_description, nrow = 1, labeller = label_wrap_gen(14)) +
  scale_color_manual(values=c('revenue loss' = "#D08770", 'revenue gain' = '#647D64'), name = "", labels = c('revenue loss' = "Loss in Revenue",  'revenue gain' = "Gain in Revenue")) +
  scale_fill_manual(values=c('revenue loss' = "#D08770", 'revenue gain' = '#647D64'), name = "", labels = c('revenue loss' = "Loss in Revenue",  'revenue gain' = "Gain in Revenue")) +
  theme(panel.grid.major.y = element_line(colour = "grey95"),
        strip.background = element_blank(), 
        axis.text = element_text(size = 14, family = "serif", color = "black"), 
        strip.text = element_text(size= 14),
        #strip.text.x = element_blank(),
        axis.title.y=element_blank(),
        axis.ticks.y=element_blank(),
        legend.position = "bottom") +
    xlab("Relative Stability Index") 

SI_ll

```

## Salmon depreciation figures

```{r, fig.width = 6, fig.height = 6}

salmon_dep <- db %>% filter(IOPAC_PORT_GROUP != "COASTWIDE" & MANAGEMENT_GROUP_CODE != "TOTAL" &
                              CRAB_YEAR %in% c(2000:2007)) %>%
                group_by(IOPAC_PORT_GROUP,CRAB_YEAR) %>% mutate(total=sum(AFI_REV_m),share=AFI_REV_m/total) %>%
                   filter(MANAGEMENT_GROUP_CODE == "SAMN") %>% 
                     group_by(IOPAC_PORT_GROUP) %>% dplyr::summarise(samn_dep=quantile(share,probs=0.5,na.rm=T)*100) %>% 
    mutate(IOPAC_PORT_GROUP_LC = str_to_title(IOPAC_PORT_GROUP),
                                IOPAC_PORT_GROUP_TT = case_when(IOPAC_PORT_GROUP_LC == "Wa Coast" ~ "WA Coast",
                                                                IOPAC_PORT_GROUP_LC == "North Wa Coast" ~ "North WA Coast",
                                                                T ~ IOPAC_PORT_GROUP_LC))
 

sens <- index_table %>% filter(Shock_id =="M2") %>% 
           ungroup() %>% select(IOPAC_PORT_GROUP,sensitivity_index)

salmon_dep_df <- salmon_dep %>% left_join(sens,by=c("IOPAC_PORT_GROUP")) %>% 
  mutate(geographic_exposure = case_when(IOPAC_PORT_GROUP %in% c('BODEGA BAY', 'MORRO', 'SAN FRANCISCO', 'MONTEREY', 'SANTA BARBARA', 'LOS ANGELES', 'SAN DIEGO') ~"High Exposure",
                                                                                                                 IOPAC_PORT_GROUP %in% c('FORT BRAGG', 'COOS BAY') ~"Medium Exposure",
                                                                                                                 IOPAC_PORT_GROUP %in% c('PUGET SOUND', 'NORTH WA COAST', 'TILLAMOOK', 'NEWPORT', 'ASTORIA', 
                                                                                                                                         'WA COAST') ~"No Exposure",
                                                                                                                 T ~ "exclude")) %>% 
  filter(geographic_exposure != 'exclude')


max_dep <- max(salmon_dep_df$samn_dep)/2


ggplot(salmon_dep_df,aes(x=samn_dep,y=sensitivity_index,label=IOPAC_PORT_GROUP_TT)) +
  geom_hline(yintercept = 0, color = "gray80") +
  geom_vline(xintercept = max_dep, color = "gray80") +
  geom_point(aes(color = geographic_exposure, shape = geographic_exposure), size = 3.25, alpha = 0.65) +
  geom_text_repel(seed = 42, size = 2.5, box.padding = 0.45, point.padding = .25, nudge_y = -.003, segment.alpha = 0.25, min.segment.length = 0, family = "sans") +
  geom_label(aes(x = 0, y = 0.05, label = "High Resilience, Low Salmon Dependence"), size = 3, hjust = "inward", family = "serif") +
  geom_label(aes(x = 20, y = 0.05, label = "High Resilience, High Salmon Dependence"), size = 3, hjust = "outward", family = "serif") +
  geom_label(aes(x = 0, y = -0.02, label = "Low Resilience, Low Salmon Dependence"), size = 3, hjust = "inward", family = "serif") +
  geom_label(aes(x = 20, y = -0.02, label = "Low Resilience, High Salmon Dependence"), size = 3, hjust = "outward", family = "serif") +
  scale_color_manual(values= nord("red_mountain", 3), name = "Geographic Exposure") +
  scale_shape_manual(values=c(17, 15, 16), name = "Geographic Exposure")+
  theme(legend.position = "bottom") +
  labs(y = 'Relative Stability Index', x = 'Salmon Dependence') 

```
